---
title: Simulated Annealing for Continuous Parameter Estimation
---

```{r libraries, include=FALSE}
library(GenSA)
library(ggplot2)
```

The quantiles given below are taken from [this PubMed article](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3946694/).

```{r}
df <- data.frame(quantile = c(0.05, 0.10, 0.25, 0.50, 0.75, 0.90, 0.95),
                 value = c(4.7, 4.8, 5.1, 5.4, 5.7, 6.1, 6.8))
df
```

We are going to assume a gamma distribution, since this allows for the skewed nature of this data (mean = 5.6, mode = 5.4).

Create a function that accepts an array of parameters. For a gamma distribution, the parameters are `shape` and `scale`. This function should return the sum squared error of difference between the actual and estimated values. The `GenSA` function will try to minimize this error.

```{r}
fun <- function (params) {
  shape <- params[1]
  scale <- params[2]
  df$est <- qgamma(df$quantile, shape, scale)
  df$sq_err <- (df$est - df$value) ^ 2
  error <- sum(df$sq_err)
  error
}
```

```{r}
lower_bound <- 1e-6
upper_bound <- 100

fit <- GenSA(par = c(1, 1), 
             fn = fun, 
             lower = c(lower_bound, lower_bound), 
             upper = c(upper_bound, upper_bound))
```

This fit yields a gamma distribution with parameters `r fit$par[1]` and `r fit$par[2]`.

We can now plot this result, comparing the estimate to the actual values.

```{r}
plot_df <- data.frame(x = seq(from = 3.0, to = 10.0, by = 0.01))
plot_df$y = pgamma(plot_df$x, fit$par[1], fit$par[2])

ggplot() +
  geom_line(aes(x, y), plot_df, color = "red") +
  geom_point(aes(value, quantile), df, size = 2, color = "blue") +
  ggtitle("Actual vs Estimate HbA1c") + 
  xlab("HbA1c") +
  ylab("Quantile")
```
